---
title: "species_area_rarefied_pooled"
output: html_document
date: "2024-06-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(purrr)
library(dplyr)
library(ggplot2)

library(MetBrewer)
library(patchwork)
library(pals)
```


```{r}
fung_data <- as.data.frame(readRDS("../intermediates/fung_rarefied_pooled_asv_table.rds"))
bact_data <- as.data.frame(readRDS("../intermediates/bact_rarefied_pooled_asv_table.rds"))

fung_rep_dat_full <- readRDS("../intermediates/fung_replication_data_v2.rds")
#fung_rep_dat <- fung_rep_dat[fung_rep_dat$sample %in% rownames(fung_data),]

bact_rep_dat_full <- readRDS("../intermediates/bact_replication_data_v2.rds")
#bact_rep_dat <- bact_rep_dat[bact_rep_dat$sample %in% rownames(bact_data),]

bro_metadata <- read.csv("../metadata/sar_metadata_clean - Sheet1.csv")
```


Table cleanup

```{r}
## fung
fung_rich <- fung_data != 0
  
fung_rep_dat <- data.frame(pooled_sample = rownames(fung_data),
                           pooled_sample2 = rownames(fung_data))
fung_rep_dat <- fung_rep_dat %>% tidyr::separate(pooled_sample2, c("bronum", "compartment"))

fung_rep_dat$richness <- rowSums(fung_rich)

fung_rep_dat_full$poolname <- paste(fung_rep_dat_full$Bro1, fung_rep_dat_full$Compartment, sep="_")
#fung_rep_dat_full <- dplyr::select(fung_rep_dat_full, -c(repnum, num))

fung_rep_dat_full <- fung_rep_dat_full %>% dplyr::group_by(poolname) %>% dplyr::mutate(repnum = length(sample)) %>% dplyr::ungroup()


fung_rep_dat$repnum <- fung_rep_dat_full$repnum[match(fung_rep_dat$pooled_sample, fung_rep_dat_full$poolname)]

fung_rep_dat_full[fung_rep_dat_full$newCompartment=="Air",] %>% dplyr::group_by(bro_and_comp) %>% dplyr::summarise(n = n())

fung_rep_dat$repnum[grepl("T1", fung_rep_dat$pooled_sample)] <- 6
fung_rep_dat$repnum[grepl("T2", fung_rep_dat$pooled_sample)] <- 6
fung_rep_dat$repnum[grepl("T3", fung_rep_dat$pooled_sample)] <- 6
fung_rep_dat$repnum[grepl("T4", fung_rep_dat$pooled_sample)] <- 5

fung_rep_dat <- fung_rep_dat %>%
  dplyr::mutate(newCompartment = case_when(
    fung_rep_dat$compartment == "I" ~ "Inverts",
    fung_rep_dat$compartment == "D" ~ "Detritus",
    grepl("T", fung_rep_dat$compartment) ~ "Air",
    grepl("W", fung_rep_dat$compartment) ~ "Water"
  ))


## bact
bact_rich <- bact_data != 0
  
bact_rep_dat <- data.frame(pooled_sample = rownames(bact_data),
                           pooled_sample2 = rownames(bact_data))
bact_rep_dat <- bact_rep_dat %>% tidyr::separate(pooled_sample2, c("bronum", "compartment"))

bact_rep_dat$richness <- rowSums(bact_rich)

bact_rep_dat_full$poolname <- paste(bact_rep_dat_full$Bro1, bact_rep_dat_full$Compartment, sep="_")
#bact_rep_dat_full <- dplyr::select(bact_rep_dat_full, -c(repnum, num))

bact_rep_dat_full <- bact_rep_dat_full %>% dplyr::group_by(poolname) %>% dplyr::mutate(repnum = length(sample)) %>% dplyr::ungroup()


bact_rep_dat$repnum <- bact_rep_dat_full$repnum[match(bact_rep_dat$pooled_sample, bact_rep_dat_full$poolname)]

bact_rep_dat_full[bact_rep_dat_full$newCompartment=="Air",] %>% dplyr::group_by(Compartment) %>% dplyr::summarise(n = n())

bact_rep_dat$repnum[grepl("T1", bact_rep_dat$pooled_sample)] <- 3
bact_rep_dat$repnum[grepl("T2", bact_rep_dat$pooled_sample)] <- 5
bact_rep_dat$repnum[grepl("T3", bact_rep_dat$pooled_sample)] <- 5
bact_rep_dat$repnum[grepl("T4", bact_rep_dat$pooled_sample)] <- 4

bact_rep_dat <- bact_rep_dat %>%
  mutate(newCompartment = case_when(
    bact_rep_dat$compartment == "I" ~ "Inverts",
    bact_rep_dat$compartment == "D" ~ "Detritus",
    grepl("T", bact_rep_dat$compartment) ~ "Air",
    grepl("W", bact_rep_dat$compartment) ~ "Water"
  ))
```
Get Hill number 1 diversity estimates for rarefied pools since they don't asymptote

```{r}

fung_inext <- readRDS("../outputs/fung_rarepool_inext_results.rds")

bact_inext <- readRDS("../outputs/bact_rarepool_inext_results.rds")


fung_asymp_shannon <- fung_inext$AsyEst[fung_inext$AsyEst$Diversity=="Shannon diversity",]

bact_asymp_shannon <- bact_inext$AsyEst[fung_inext$AsyEst$Diversity=="Shannon diversity",]


fung_rep_dat$shannon <- fung_asymp_shannon$Estimator[match(fung_asymp_shannon$Site, fung_rep_dat$pooled_sample)]

bact_rep_dat$shannon <- bact_asymp_shannon$Estimator[match(bact_asymp_shannon$Site, bact_rep_dat$pooled_sample)]


fung_rep_dat$log_shannon <- log(fung_rep_dat$shannon)
bact_rep_dat$log_shannon <- log(bact_rep_dat$shannon)
```

```{r}
brostuff <- bro_metadata %>% tidyr::separate(Label, c("Project", "bro_num", "num"))
brostuff$bro_num <- paste0(brostuff$bro_num, brostuff$num)

fung_rep_dat$water_volume <- brostuff$Water.volume...in.lab...mL.[match(fung_rep_dat$bronum, brostuff$bro_num)]
fung_rep_dat$detritus_wt <- brostuff$Dry.weight.detritus[match(fung_rep_dat$bronum, brostuff$bro_num)]

bact_rep_dat$water_volume <- brostuff$Water.volume...in.lab...mL.[match(bact_rep_dat$bronum, brostuff$bro_num)]
bact_rep_dat$detritus_wt <- brostuff$Dry.weight.detritus[match(bact_rep_dat$bronum, brostuff$bro_num)]
```

```{r}
# fung_rep_dat$newCompartment[fung_rep_dat$Compartment=="W02"] = "W02"
# fung_rep_dat$newCompartment[fung_rep_dat$Compartment=="WPool"] = "WPool"

# remove Wpool samples in bact?
bact_rep_dat <- bact_rep_dat[!(grepl("WPool", bact_rep_dat$pooled_sample)),]

bact_data <- bact_data[rownames(bact_data) %in% bact_rep_dat$pooled_sample,]
bact_data <- bact_data[,colSums(bact_data)>0]
```

Functions!
```{r}
get_data_for_regress <- function(rep_dat, compartment) {
  filt_dat <- rep_dat[rep_dat$newCompartment == compartment,]
  
  #filt_dat$log_shannon <- log(filt_dat$richness)
  
  # dat_summ <- filt_dat %>% group_by(bronum) %>% summarize(log_rich = mean(log_shannon))
  # 
  # dat_summ_se <- filt_dat %>% group_by(bronum) %>% 
  #   summarize(se = sd(log_shannon)/sqrt(length(log_shannon)))
  # 
  # dat_summ$se <- dat_summ_se$se
  # 
  # dat_summ$water_volume <- filt_dat$water_volume[match(dat_summ$bronum, filt_dat$bronum)]
  # dat_summ$detritus_wt <- filt_dat$detritus_wt[match(dat_summ$bronum, filt_dat$bronum)]
  
  # res <- list()
  
  # res[["full_data"]] <- filt_dat
  # res[["summarized_data"]] <- dat_summ
  
  return(filt_dat)
}

fung_inv_regress_dat <- get_data_for_regress(rep_dat = fung_rep_dat,
                                             compartment = "Inverts")

fung_det_regress_dat <- get_data_for_regress(rep_dat = fung_rep_dat,
                                             compartment = "Detritus")

fung_water_regress_dat <- get_data_for_regress(rep_dat = fung_rep_dat,
                                             compartment = "Water")




bact_inv_regress_dat <- get_data_for_regress(rep_dat = bact_rep_dat,
                                             compartment = "Inverts")

bact_det_regress_dat <- get_data_for_regress(rep_dat = bact_rep_dat,
                                             compartment = "Detritus")

bact_water_regress_dat <- get_data_for_regress(rep_dat = bact_rep_dat,
                                             compartment = "Water")



```


```{r}
# bootstrap function
bfun <- \(x, y, R=1000, seed=134) {
  FUN <- \() {
    i <- sample(seq_len(length(x)), replace=TRUE)
    y <- y[i]
    x <- x[i]
    return(lm(y ~ x)$coefficients)

    #return()
  }
  set.seed(seed)
  r <- t(replicate(R, FUN()))
  bootdist <- r[, 2]

  return(bootdist)

}

```

Regressions and bootstraps
```{r}
# volume = detritus weight
fung_regress_data_list = list(fung_inv_regress_dat,
                              fung_det_regress_dat,
                              fung_water_regress_dat)

fung_detwt_boot_slopes <- lapply(fung_regress_data_list,
                                 function(data) bfun(x = log(data$detritus_wt), y = data$log_shannon, R = 10000, seed=1899))


fung_regular_lms_detritus_wt <- lapply(fung_regress_data_list,
                           function(x) lm(log_shannon ~ log(detritus_wt), data = x))

fung_reg_lm_slopes_detritus_wt <- lapply(fung_regular_lms_detritus_wt, function(y) coef(y)[2])

fung_tts_detritus_wt <- list()

for (i in 1:length(fung_regress_data_list)) {
  fung_tts_detritus_wt[[i]] = t.test(unlist(fung_detwt_boot_slopes[[i]]), mu = fung_reg_lm_slopes_detritus_wt[[i]])
}




##### water volume
fung_wvol_boot_slopes <- lapply(fung_regress_data_list,
                                 function(data) bfun(x = log(data$water_volume), y = data$log_shannon, R = 10000, seed=222))

fung_regular_lms_water_volume <- lapply(fung_regress_data_list,
                           function(x) lm(log_shannon ~ log(water_volume), data = x))

fung_reg_lm_slopes_water_volume <- lapply(fung_regular_lms_water_volume, function(y) coef(y)[2])

fung_tts_water_volume <- list()

for (i in 1:length(fung_regress_data_list)) {
  fung_tts_water_volume[[i]] = t.test(unlist(fung_wvol_boot_slopes[[i]]), mu = fung_reg_lm_slopes_water_volume[[i]])
}
```

Air
```{r}
# air
# richness per sample vs water volume

# We left the sampler out in the field to collect samples of the air microbiome above the bromeliad garden for 1 hour, 2.5 hours, 4.5 hours, and 6 hours with n=6 replicates per time point. These time points correspond to 990, 2475, 4455, and 5940 liters respectively. 

fung_air <- fung_rep_dat[fung_rep_dat$newCompartment=="Air",]
fung_air <- subset(fung_air, select = -c(detritus_wt, water_volume))
fung_air$combo <- paste(fung_air$bronum, fung_air$compartment, sep = "_")

fung_air<- fung_air %>%
  mutate(air_volume = case_when(
    compartment == "T1" ~ 990,
    compartment == "T2" ~ 2475,
    compartment == "T3" ~ 4455,
    compartment == "T4" ~ 5940
  ))


##### bootstrap and regressions
fung_air_boot_slopes <- bfun(x = log(fung_air$air_volume), y = fung_air$log_shannon, R = 10000, seed=111)

fung_regular_lm_air_volume <- lm(log_shannon ~ log(air_volume), data = fung_air)

fung_reg_lm_slope_air_volume <- coef(fung_regular_lm_air_volume)[2]

fung_tts_air_volume <- t.test(unlist(fung_air_boot_slopes),
                              mu = fung_reg_lm_slope_air_volume)


```

Plots 
```{r}
my_regress_plot <- function(regress_data, wvol_color = "black", detwt_color = "black") {
  gg1 <- ggplot(regress_data, aes(x=log(water_volume), y=log_shannon)) + 
  geom_point(color = wvol_color) +
  stat_smooth(method = "lm", 
              formula = y ~ x, 
              geom = "smooth",
              se = F,
              color = wvol_color) +
  #theme(axis.text = element_text(color = "black", size = 14)) +
  labs(y = "log(Shannon)") +
  theme_classic() +
  theme(axis.text = element_text(color = "black", size = 12),
        axis.title = element_text(color = "black", size = 13)) 
  
  gg2 <- ggplot(regress_data, aes(x=log(detritus_wt), y=log_shannon)) + 
  geom_point(color = detwt_color) +
  stat_smooth(method = "lm", 
              formula = y ~ x, 
              geom = "smooth",
              se = F,
              color = detwt_color) +
  # geom_errorbar(aes(ymin=log_rich-se, ymax=log_rich+se),
  #               width=.2,
  #               color = met.brewer("NewKingdom")[4]) +
  labs(y = "log(Shannon)", x = "log(detritus_weight)") +
  theme_classic() +
  theme(axis.text = element_text(color = "black", size = 12),
        axis.title = element_text(color = "black", size = 13)) 
  
  gglist <- list(gg1, gg2)
  
  return(gglist)
  
  
}


fung_inv_plot <- my_regress_plot(fung_inv_regress_dat, 
                                 wvol_color = stepped()[1],
                                 detwt_color = stepped()[5])
fung_inv_plot <- lapply(fung_inv_plot, function(x) x + ggtitle("Fungi, Invertebrates"))

fung_det_plot <- my_regress_plot(fung_det_regress_dat,
                                 detwt_color = stepped()[9])
fung_det_plot <- lapply(fung_det_plot, function(x) x + ggtitle("Fungi, Detritus"))

fung_water_plot <- my_regress_plot(fung_water_regress_dat,
                                   wvol_color = stepped()[13])
fung_water_plot <- lapply(fung_water_plot, function(x) x + ggtitle("Fungi, Water"))

# fung_wpool_plot <- my_regress_plot(fung_wpool_regress_dat$summarized_data)
# fung_wpool_plot <- lapply(fung_wpool_plot, function(x) x + ggtitle("Fungi, WPool"))

fung_air_plot <- ggplot(fung_air, aes(x=log(air_volume), y=log_shannon)) + 
  geom_point(color = stepped()[19]) +
  stat_smooth(method = "lm", 
              formula = y ~ x, 
              geom = "smooth",
              se = F,
              color = stepped()[19]) +
  # geom_errorbar(aes(ymin=log_rich-se, ymax=log_rich+se), 
  #               width=.2,
  #               color = met.brewer("NewKingdom")[1]) +
  labs(y = "log(Shannon)", title = "Fungi, Air") +
  theme_classic() +
  theme(axis.text = element_text(color = "black", size = 12),
        axis.title = element_text(color = "black", size = 13)) 

fung_plots <- fung_inv_plot[[1]] + fung_inv_plot[[2]] +
  fung_water_plot[[1]] +
  fung_det_plot [[2]] + plot_layout(nrow = 2)


#ggsave("../figures/rarepool_all_fungi_sar_regression_plots.png", fung_plots, height = 12, width = 9.5)

```


Table
```{r}
# things to put in table = locus, compartment, R2, p-value, slope, t-test p-value, intercept

fung_tbl_detritus_wt <- data.frame(Locus = rep("Fungi", 3),
                       Compartment = c("Invertebrates", "Detritus", "Water (W02 and WPool)"),
                       Volume = rep("Detritus_weight_mg", 3))

fung_tbl_detritus_wt$R2 = sapply(fung_regular_lms_detritus_wt,
                     function(x) summary(x)$r.squared)

fung_tbl_detritus_wt$pval = sapply(fung_regular_lms_detritus_wt,
                       function(x) as.numeric(summary(x)$coefficients[,4][2]))

fung_tbl_detritus_wt$Slope = sapply(fung_regular_lms_detritus_wt,
                        function(x) coef(x)[2])

fung_tbl_detritus_wt$bootstrap_ttest_pval <- sapply(fung_tts_detritus_wt, 
                                        function(x) x$p.value)

fung_tbl_detritus_wt$Intercept = sapply(fung_regular_lms_detritus_wt,
                        function(x) coef(x)[1])


##### water volume
fung_tbl_water_volume <- data.frame(Locus = rep("Fungi", 3),
                       Compartment = c("Invertebrates", "Detritus", "Water (W02 and WPool)"),
                       Volume = rep("Water_volume_mL", 3))

fung_tbl_water_volume$R2 = sapply(fung_regular_lms_water_volume,
                     function(x) summary(x)$r.squared)

fung_tbl_water_volume$pval = sapply(fung_regular_lms_water_volume,
                       function(x) as.numeric(summary(x)$coefficients[,4][2]))

fung_tbl_water_volume$Slope = sapply(fung_regular_lms_water_volume,
                        function(x) coef(x)[2])

fung_tbl_water_volume$bootstrap_ttest_pval <- sapply(fung_tts_water_volume, 
                                        function(x) x$p.value)

fung_tbl_water_volume$Intercept = sapply(fung_regular_lms_water_volume,
                        function(x) coef(x)[1])


fung_full_tbl <- rbind(fung_tbl_detritus_wt, fung_tbl_water_volume)


fung_row_air <- c(Locus = "Fungi",
                  Compartment = "Air",
                  Volume = "Air_volume_L",
                  R2 = summary(fung_regular_lm_air_volume)$r.squared,
                  pval = as.numeric(summary(fung_regular_lm_air_volume)$coefficients[,4][2]),
                  Slope = coef(fung_regular_lm_air_volume)[2],
                  bootstrap_ttest_pval = fung_tts_air_volume$p.value,
                  Intercept = coef(fung_regular_lm_air_volume)[1]
                  )

fung_full_tbl <- rbind(fung_full_tbl, fung_row_air)

# take out things not focusing on
fung_final_tbl <- fung_full_tbl[-c(5,3),]

fung_final_tbl <- fung_final_tbl[c(3,1,4,2,5),]

```


# Bacteria

Regressions and bootstraps
```{r}
# volume = detritus weight
bact_regress_data_list = list(bact_inv_regress_dat,
                              bact_det_regress_dat,
                              bact_water_regress_dat)

bact_detwt_boot_slopes <- lapply(bact_regress_data_list,
                                 function(data) bfun(x = log(data$detritus_wt), y = data$log_shannon, R = 10000, seed=18991))


bact_regular_lms_detritus_wt <- lapply(bact_regress_data_list,
                           function(x) lm(log_shannon ~ log(detritus_wt), data = x))

bact_reg_lm_slopes_detritus_wt <- lapply(bact_regular_lms_detritus_wt, function(y) coef(y)[2])

bact_tts_detritus_wt <- list()

for (i in 1:length(bact_regress_data_list)) {
  bact_tts_detritus_wt[[i]] = t.test(unlist(bact_detwt_boot_slopes[[i]]), mu = bact_reg_lm_slopes_detritus_wt[[i]])
}




##### water volume
bact_wvol_boot_slopes <- lapply(bact_regress_data_list,
                                 function(data) bfun(x = log(data$water_volume), y = data$log_shannon, R = 10000, seed=2221))

bact_regular_lms_water_volume <- lapply(bact_regress_data_list,
                           function(x) lm(log_shannon ~ log(water_volume), data = x))

bact_reg_lm_slopes_water_volume <- lapply(bact_regular_lms_water_volume, function(y) coef(y)[2])

bact_tts_water_volume <- list()

for (i in 1:length(bact_regress_data_list)) {
  bact_tts_water_volume[[i]] = t.test(unlist(bact_wvol_boot_slopes[[i]]), mu = bact_reg_lm_slopes_water_volume[[i]])
}
```

Air
```{r}
# air
# richness per sample vs water volume

# We left the sampler out in the field to collect samples of the air microbiome above the bromeliad garden for 1 hour, 2.5 hours, 4.5 hours, and 6 hours with n=6 replicates per time point. These time points correspond to 990, 2475, 4455, and 5940 liters respectively. 

bact_air <- bact_rep_dat[bact_rep_dat$newCompartment=="Air",]
bact_air <- subset(bact_air, select = -c(detritus_wt, water_volume))
bact_air$combo <- paste(bact_air$bronum, bact_air$compartment, sep = "_")

bact_air<- bact_air %>%
  mutate(air_volume = case_when(
    compartment == "T1" ~ 990,
    compartment == "T2" ~ 2475,
    compartment == "T3" ~ 4455,
    compartment == "T4" ~ 5940
  ))


##### bootstrap and regressions
bact_air_boot_slopes <- bfun(x = log(bact_air$air_volume), y = bact_air$log_shannon, R = 10000, seed=1111)

bact_regular_lm_air_volume <- lm(log_shannon ~ log(air_volume), data = bact_air)

bact_reg_lm_slope_air_volume <- coef(bact_regular_lm_air_volume)[2]

bact_tts_air_volume <- t.test(unlist(bact_air_boot_slopes),
                              mu = bact_reg_lm_slope_air_volume)


```

Plots 
```{r}

bact_inv_plot <- my_regress_plot(bact_inv_regress_dat, 
                                 wvol_color = stepped()[1],
                                 detwt_color = stepped()[5])
bact_inv_plot <- lapply(bact_inv_plot, function(x) x + ggtitle("Bacteria, Invertebrates"))

bact_det_plot <- my_regress_plot(bact_det_regress_dat,
                                 detwt_color = stepped()[9])
bact_det_plot <- lapply(bact_det_plot, function(x) x + ggtitle("Bacteria, Detritus"))

bact_water_plot <- my_regress_plot(bact_water_regress_dat,
                                   wvol_color = stepped()[13])
bact_water_plot <- lapply(bact_water_plot, function(x) x + ggtitle("Bacteria, Water"))

# bact_wpool_plot <- my_regress_plot(bact_wpool_regress_dat$summarized_data)
# bact_wpool_plot <- lapply(bact_wpool_plot, function(x) x + ggtitle("Bacteria, WPool"))

bact_air_plot <- ggplot(bact_air, aes(x=log(air_volume), y=log_shannon)) + 
  geom_point(color = stepped()[19]) +
  stat_smooth(method = "lm", 
              formula = y ~ x, 
              geom = "smooth",
              se = F,
              color = stepped()[19]) +
  # geom_errorbar(aes(ymin=log_rich-se, ymax=log_rich+se), 
  #               width=.2,
  #               color = met.brewer("NewKingdom")[1]) +
  labs(y = "log(Shannon)", title = "Bacteria, Air") +
  theme_classic() +
  theme(axis.text = element_text(color = "black", size = 12),
        axis.title = element_text(color = "black", size = 13)) 

bact_plots <- bact_inv_plot[[1]] + bact_inv_plot[[2]] +
  bact_water_plot[[1]] +
  bact_det_plot [[2]] + plot_layout(nrow = 2)


all_sar_plots <- 
  fung_det_plot [[2]] + fung_water_plot[[1]] + fung_inv_plot[[2]] + fung_inv_plot[[1]] + 
  bact_det_plot [[2]] + bact_water_plot[[1]] + bact_inv_plot[[2]] + bact_inv_plot[[1]] +
  plot_layout(nrow = 2)

ggsave("../figures/for_manuscript/all_sar_rarepool_plots.pdf", width=16, height=7.5)
ggsave("../figures/for_manuscript/all_sar_rarepool_plots.png", width=16, height=7.5)


#ggsave("../figures/rarepool_all_bact_sar_regression_plots.png", bact_plots, height = 12, width = 9.5)

```

Table
```{r}
# things to put in table = locus, compartment, R2, p-value, slope, t-test p-value, intercept

bact_tbl_detritus_wt <- data.frame(Locus = rep("Bacteria", 3),
                       Compartment = c("Invertebrates", "Detritus", "Water (W02)"),
                       Volume = rep("Detritus_weight_mg", 3))

bact_tbl_detritus_wt$R2 = sapply(bact_regular_lms_detritus_wt,
                     function(x) summary(x)$r.squared)

bact_tbl_detritus_wt$pval = sapply(bact_regular_lms_detritus_wt,
                       function(x) as.numeric(summary(x)$coefficients[,4][2]))

bact_tbl_detritus_wt$Slope = sapply(bact_regular_lms_detritus_wt,
                        function(x) coef(x)[2])

bact_tbl_detritus_wt$bootstrap_ttest_pval <- sapply(bact_tts_detritus_wt, 
                                        function(x) x$p.value)

bact_tbl_detritus_wt$Intercept = sapply(bact_regular_lms_detritus_wt,
                        function(x) coef(x)[1])


##### water volume
bact_tbl_water_volume <- data.frame(Locus = rep("Bacteria", 3),
                       Compartment = c("Invertebrates", "Detritus", "Water (W02)"),
                       Volume = rep("Water_volume_mL", 3))

bact_tbl_water_volume$R2 = sapply(bact_regular_lms_water_volume,
                     function(x) summary(x)$r.squared)

bact_tbl_water_volume$pval = sapply(bact_regular_lms_water_volume,
                       function(x) as.numeric(summary(x)$coefficients[,4][2]))

bact_tbl_water_volume$Slope = sapply(bact_regular_lms_water_volume,
                        function(x) coef(x)[2])

bact_tbl_water_volume$bootstrap_ttest_pval <- sapply(bact_tts_water_volume, 
                                        function(x) x$p.value)

bact_tbl_water_volume$Intercept = sapply(bact_regular_lms_water_volume,
                        function(x) coef(x)[1])


bact_full_tbl <- rbind(bact_tbl_detritus_wt, bact_tbl_water_volume)


bact_row_air <- c(Locus = "Bacteria",
                  Compartment = "Air",
                  Volume = "Air_volume_L",
                  R2 = summary(bact_regular_lm_air_volume)$r.squared,
                  pval = as.numeric(summary(bact_regular_lm_air_volume)$coefficients[,4][2]),
                  Slope = coef(bact_regular_lm_air_volume)[2],
                  bootstrap_ttest_pval = bact_tts_air_volume$p.value,
                  Intercept = coef(bact_regular_lm_air_volume)[1]
                  )

bact_full_tbl <- rbind(bact_full_tbl, bact_row_air)

bact_final_tbl <- bact_full_tbl[-c(3,5),]
bact_final_tbl <- bact_final_tbl[c(3,1,4,2,5),]

```

```{r}
full_full_tbl <- rbind(fung_final_tbl, bact_final_tbl)
full_full_tbl$bootstrap_ttest_pval <- as.numeric(full_full_tbl$bootstrap_ttest_pval)

### correcting P-values for multiple comparisons
library(multtest)  

# Set seed for repeatability
set.seed(2024)
  
adj_p_values = mt.rawp2adjp(full_full_tbl$bootstrap_ttest_pval)
    
adj_p_mat = adj_p_values$adjp[, c("rawp", "BH")]
  
adj_p_ordered = adj_p_mat[,2][order(adj_p_values$index)]

full_full_tbl$bootstrap_ttest_pval <- adj_p_ordered

saveRDS(full_full_tbl, "../intermediates/rarepool_regression_stats_table.rds")
```


# z-value comparisons among compartments, within and between loci
Bootstrap to generate variance on the observed data slope, repeated 1,000 times
```{r}

get_ci <- function(slope_distribution_vector) {
  sample.mean <- mean(slope_distribution_vector)
  sample.n <- length(slope_distribution_vector)
  sample.sd <- sd(slope_distribution_vector)
  sample.se <- sample.sd/sqrt(sample.n)
  
  alpha = 0.05
  degrees.freedom = sample.n - 1
  t.score = qt(p=alpha/2, df=degrees.freedom,lower.tail=F)
  
  margin.error <- t.score * sample.se
  
  lower.bound <- sample.mean - margin.error
  upper.bound <- sample.mean + margin.error
  
  return(paste(round(lower.bound, 4),
               round(upper.bound, 4),
               sep=","))

}

```

```{r}
fung_dw <- lapply(fung_regress_data_list,
                  function(x) bfun(y = x$log_shannon, x = log(x$detritus_wt)))

fung_dw_ci <- lapply(fung_dw, get_ci)

fung_wv <- lapply(fung_regress_data_list,
                  function(x) bfun(y = x$log_shannon, x = log(x$water_volume)))

fung_wv_ci <- lapply(fung_wv, get_ci)

fung_a <- bfun(y = fung_air$log_shannon, x = log(fung_air$air_volume))
fung_a <- fung_a[!is.na(fung_a)]

fung_a_ci <- get_ci(fung_a)


bact_dw <- lapply(bact_regress_data_list,
                  function(x) bfun(y = x$log_shannon, x = log(x$detritus_wt)))

bact_dw_ci <- lapply(bact_dw, get_ci)

bact_wv <- lapply(bact_regress_data_list ,
                  function(x) bfun(y = x$log_shannon, x = log(x$water_volume)))

bact_wv_ci <- lapply(bact_wv, get_ci)

bact_a <- bfun(y = bact_air$log_shannon, x = log(bact_air$air_volume))
bact_a <- bact_a[!is.na(bact_a)]

bact_a_ci <- get_ci(bact_a)


# append to full table
ci_vector <- c(fung_wv_ci[[1]],
                    fung_dw_ci[[1]],
                    fung_wv_ci[[3]],
                    fung_dw_ci[[2]],
                    fung_a_ci,
                    bact_wv_ci[[1]],
                    bact_dw_ci[[1]],
                    bact_wv_ci[[3]],
                    bact_dw_ci[[2]],
                    bact_a_ci
                    )

new_full_table <- full_full_tbl
new_full_table$CI_95 <- ci_vector

new_full_table <- new_full_table %>% dplyr::select(Locus, Compartment, Volume, R2, pval, Slope, CI_95, bootstrap_ttest_pval, Intercept)
saveRDS(new_full_table, "../intermediates/rarepool_regression_stats_with_bootstrapped_confint.rds")

write.csv(new_full_table, "../outputs/rarepool_regression_stats_with_bootstrapped_confint.csv")
```

Pairwise t-tests
```{r}
# inv, det, water, air
long_data <- data.frame(type = c(rep("fung_inverts_dw", length(fung_dw[[1]])),
                                 rep("fung_detritus_dw", length(fung_dw[[2]])),
                                 #rep("fung_water_dw", length(fung_dw[[3]])),
                                 rep("fung_inverts_wv", length(fung_wv[[1]])),
                                 #rep("fung_detritus_wv", length(fung_wv[[2]])),
                                 rep("fung_water_wv", length(fung_wv[[3]])),
                                 rep("fung_air", length(fung_a)),
                                 
                                 rep("bact_inverts_dw", length(bact_dw[[1]])),
                                 rep("bact_detritus_dw", length(bact_dw[[2]])),
                                 #rep("bact_water_dw", length(bact_dw[[3]])),
                                 rep("bact_inverts_wv", length(bact_wv[[1]])),
                                 #rep("bact_detritus_wv", length(bact_wv[[2]])),
                                 rep("bact_water_wv", length(bact_wv[[3]])),
                                 rep("bact_air", length(bact_a))
                                 
                                 ),
                        slope = c(fung_dw[[1]],
                                  fung_dw[[2]],
                                  #fung_dw[[3]],
                                  fung_wv[[1]],
                                  #fung_wv[[2]],
                                  fung_wv[[3]],
                                  fung_a,
                                  
                                  bact_dw[[1]],
                                  bact_dw[[2]],
                                  #bact_dw[[3]],
                                  bact_wv[[1]],
                                  #bact_wv[[2]],
                                  bact_wv[[3]],
                                  bact_a
                                  ))


pair_t <- pairwise.t.test(long_data$slope, long_data$type, p.adjust.method = "BH")[[3]]
pair_t2 <- apply(pair_t, 2, function(x) ifelse(!is.na(x), round(x, 7), NA))

saveRDS(pair_t2, "../intermediates/rarepool_pairwise_ttest_boot_slopes_between_compartments.rds")

write.csv(pair_t2, "../outputs/rarepool_pairwise_ttest_boot_slopes_between_compartments.csv")
```

## save out data for slope comparison
```{r}
saveRDS(fung_regress_data_list, "../intermediates/rarepool_fung_regress_data_list.rds")
saveRDS(fung_air, "../intermediates/rarepool_fung_air_data.rds")

saveRDS(bact_regress_data_list, "../intermediates/rarepool_bact_regress_data_list.rds")
saveRDS(bact_air, "../intermediates/rarepool_bact_air_data.rds")
```

Some regressions
```{r}

gg <- ggplot(fung_inv_summ, aes(x=log(water_vol), y=log(rich))) + geom_point()

summary(inv_wat1 <- lm(rich ~ water_vol, data = fung_inv_summ))
summary(inv_wat2<- lm(log(rich) ~ log(water_vol), data = fung_inv_summ))
summary(inv_wat3 <- glm.nb(rich ~ water_vol, data = fung_inv_summ))

AIC(inv_wat1)
AIC(inv_wat2)
AIC(inv_wat3)

summary(inv_det1 <- lm(rich ~ detritus_wt, data = fung_inv_summ))
summary(inv_det2<- lm(log(rich) ~ log(detritus_wt), data = fung_inv_summ))
summary(inv_det3 <- glm.nb(rich ~ detritus_wt, data = fung_inv_summ))

AIC(inv_det1)
AIC(inv_det2)
AIC(inv_det3)


# test <- lm(log(fung_rep_dat$richness) ~ log(fung_rep_dat$water_volume))
# summary(test1 <- glm.nb(richness ~ water_volume, data=fung_rep_dat))
```



```{r}
#gg <- ggplot(fung_air_summ, aes(x=log(water_vol), y=log(rich))) + geom_point()

summary(det_wat1 <- lm(rich ~ water_vol, data = fung_det_summ))
summary(det_wat2<- lm(log(rich) ~ log(water_vol), data = fung_det_summ))
#summary(det_wat3 <- glm.nb(rich ~ water_vol, data = fung_det_summ))

AIC(det_wat1)
AIC(det_wat2)
#AIC(det_wat3)

summary(det_det1 <- lm(rich ~ detritus_wt, data = fung_det_summ))
summary(det_det2<- lm(log(rich) ~ log(detritus_wt), data = fung_det_summ))
#summary(det_det3 <- glm.nb(rich ~ detritus_wt, data = fung_det_summ))

AIC(det_det1)
AIC(det_det2)
#AIC(det_det3)

```

