---
title: "metadata_multiple_regression"
author: "Kacie Kajihara"
date: "2024-06-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(relaimpo)
library(dplyr)
library(tibble)
library(ggplot2)
library(stargazer)
```


```{r}
bro_metadata <- read.csv("../metadata/sar_metadata_clean - Sheet1.csv")

brostuff <- bro_metadata %>% tidyr::separate(Label, c("Project", "bro_num", "num"))
brostuff$bro_num <- paste0(brostuff$bro_num, brostuff$num)

fung_regress_dat_list <- readRDS("../intermediates/truly_nonpooled_fung_regress_data_list.rds")

fung_inv_regress_dat <- fung_regress_dat_list[[1]]
fung_det_regress_dat <- fung_regress_dat_list[[2]]
fung_water_regress_dat <- fung_regress_dat_list[[3]]

bact_regress_dat_list <- readRDS("../intermediates/truly_nonpooled_bact_regress_data_list.rds")

bact_inv_regress_dat <- bact_regress_dat_list[[1]]
bact_det_regress_dat <- bact_regress_dat_list[[2]]
bact_water_regress_dat <- bact_regress_dat_list[[3]]

```

```{r}
detritus_wt_meta <- brostuff %>% dplyr::select(bro_num,
                                            Diameter,
                                            Height,
                                            Complexity,
                                            Number.of..invertebrates,
                                            X.C..detritus.,
                                            X.N..detritus.,
                                            Invert.Weight)

names(detritus_wt_meta) <- c("bro_num", "Diameter", "Height", "Complexity", "Invertebrate_number",
                            "pctC_detritus", "pctN_detritus", "Invertebrate_weight")

#detritus_wt_meta <- column_to_rownames(detritus_wt_meta, "bro_num")

water_vol_meta <- brostuff %>% dplyr::select(bro_num,
                                            Diameter,
                                            Height,
                                            Complexity,
                                            Number.of..invertebrates,
                                            Dissolved.Oyxgen,
                                            pH,
                                            Temperature,
                                            Nitrate..Water.,
                                            Invert.Weight)

names(water_vol_meta) <- c("bro_num", "Diameter", "Height", "Complexity", "Invertebrate_number",
                            "Dissolved_oxygen", "pH", "Temperature", "Nitrate_water", "Invertebrate_weight")

#water_vol_meta <- column_to_rownames(water_vol_meta, "bro_num")
```

```{r}
# dw_gg <- GGally::ggpairs(detritus_wt_meta[,-1]) + theme_bw()
# 
# wv_gg <- GGally::ggpairs(water_vol_meta[,-1]) + theme_bw()
```

```{r}
append_meta <- function(regress_data, metadata) {
  num = ncol(regress_data)
  
  for (i in 1:ncol(metadata)) {
    regress_data[,num+i] = metadata[,i][match(regress_data$Bro1, metadata$bro_num)]
    names(regress_data)[num+i] = names(metadata)[i]
  }
  
  regress_data <- dplyr::select(regress_data, -c(Bro1, repnum))
  return(regress_data)
}

# test = append_meta(fung_det_regress_dat, detritus_wt_meta)
# 
# testlm <- lm(log_richness ~ log(detritus_wt) + Diameter + Height + Complexity + Invertebrate_number +
#              + pctC_detritus + pctN_detritus, data = test)
# 
# testlm <- lm(log_richness ~ log(detritus_wt) + Diameter + Height , data = test)
# 
# hm <- calc.relimp(testlm)
# 
# wha <- stepAIC(testlm, direction="both")
```

### Fungi metadata multiple regression
```{r, echo=F, message=F, results='hide'}
# detritus weight
fung_detwt_list <- list(fung_inv_regress_dat,
                      fung_det_regress_dat)

fung_detwt_dat_meta <- fung_detwt_lms <- fung_detwt_relimp <- fung_detwt_aicmod <- list()
  
for (i in 1:length(fung_detwt_list)) {
  
  fung_detwt_dat_meta[[i]] <- append_meta(fung_detwt_list[[i]], detritus_wt_meta)
  
  fung_detwt_dat_meta_nona <- lapply(fung_detwt_dat_meta, na.exclude)

  fung_detwt_lms[[i]] <- lm(log_richness ~ log(Invertebrate_weight) 
                              + Diameter + Height
                         + Complexity + Invertebrate_number + pctC_detritus + pctN_detritus, 
                         data = fung_detwt_dat_meta_nona[[i]])
  
  fung_detwt_relimp[[i]] <- calc.relimp(fung_detwt_lms[[i]])
  
  fung_detwt_aicmod[[i]] <- holding <- stepAIC(fung_detwt_lms[[i]])

}



# water volume
fung_wvol_list <- list(fung_inv_regress_dat,
                      fung_water_regress_dat)

fung_wvol_dat_meta <- fung_wvol_lms <- fung_wvol_relimp <- fung_wvol_aicmod <- list()
  
for (i in 1:length(fung_wvol_list)) {
  
  fung_wvol_dat_meta[[i]] <- append_meta(fung_wvol_list[[i]], water_vol_meta)
  
  fung_wvol_dat_meta_nona <- lapply(fung_wvol_dat_meta, na.exclude)

  fung_wvol_lms[[i]] <- lm(log_richness ~ log(Invertebrate_weight) + Diameter + Height
                         + Complexity + Invertebrate_number + Dissolved_oxygen + pH + Temperature + Nitrate_water,
                         data = fung_wvol_dat_meta_nona[[i]])
  
  fung_wvol_relimp[[i]] <- calc.relimp(fung_wvol_lms[[i]])
  
  fung_wvol_aicmod[[i]] <- stepAIC(fung_wvol_lms[[i]])
}


```

#### Inverts - detritus weight

Model summary
```{r, echo=F}
summary(fung_detwt_lms[[1]])
```

Relative importance of each predictor
```{r, echo=F}
fung_detwt_relimp[[1]]
```

Stepwise AIC selected model
```{r, echo=F}
summary(fung_detwt_aicmod[[1]])
```

#### Detritus - detritus weight

Model summary
```{r, echo=F}
summary(fung_detwt_lms[[2]])
```

Relative importance of each predictor
```{r, echo=F}
fung_detwt_relimp[[2]]
```

Stepwise AIC selected model
```{r, echo=F}
summary(fung_detwt_aicmod[[2]])
```

#### Inverts - water volume

Model summary
```{r, echo=F}
summary(fung_wvol_lms[[1]])
```

Relative importance of each predictor
```{r, echo=F}
fung_wvol_relimp[[1]]
```

Stepwise AIC selected model
```{r, echo=F}
summary(fung_wvol_aicmod[[1]])
```

#### Water filters - water volume

Model summary
```{r, echo=F}
summary(fung_wvol_lms[[2]])
```

Relative importance of each predictor
```{r, echo=F}
fung_wvol_relimp[[2]]
```

Stepwise AIC selected model
```{r, echo=F}
summary(fung_wvol_aicmod[[2]])
```


### Bacteria metadata multiple regression
```{r, echo=F, message=F, results='hide'}
# detritus weight
bact_detwt_list <- list(bact_inv_regress_dat,
                      bact_det_regress_dat)

bact_detwt_dat_meta <- bact_detwt_lms <- bact_detwt_relimp <- bact_detwt_aicmod <- list()

# bro9 was removed bc no invert weight
  
for (i in 1:length(bact_detwt_list)) {
  
  bact_detwt_dat_meta[[i]] <- append_meta(bact_detwt_list[[i]], detritus_wt_meta)
  bact_detwt_dat_meta[[i]] <- na.omit(bact_detwt_dat_meta[[i]]) 

  bact_detwt_lms[[i]] <- lm(log_richness ~ log(Invertebrate_weight) + Diameter + Height
                         + Complexity + Invertebrate_number + pctC_detritus + pctN_detritus, 
                         data = bact_detwt_dat_meta[[i]])
  
  bact_detwt_relimp[[i]] <- calc.relimp(bact_detwt_lms[[i]])
  
  bact_detwt_aicmod[[i]] <- stepAIC(bact_detwt_lms[[i]])
}



# water volume
bact_wvol_list <- list(bact_inv_regress_dat,
                      bact_water_regress_dat)

bact_wvol_dat_meta <- bact_wvol_lms <- bact_wvol_relimp <- bact_wvol_aicmod <- list()
  
for (i in 1:length(bact_wvol_list)) {
  
  bact_wvol_dat_meta[[i]] <- append_meta(bact_wvol_list[[i]], water_vol_meta)
  
  bact_wvol_dat_meta_nona <- lapply(bact_wvol_dat_meta, na.exclude)

  bact_wvol_lms[[i]] <- lm(log_richness ~ log(Invertebrate_weight) + Diameter + Height
                         + Complexity + Invertebrate_number + Dissolved_oxygen + pH + Temperature + Nitrate_water,
                         data = bact_wvol_dat_meta_nona[[i]])
  
  bact_wvol_relimp[[i]] <- calc.relimp(bact_wvol_lms[[i]])
  
  bact_wvol_aicmod[[i]] <- stepAIC(bact_wvol_lms[[i ]])
}

```

#### Inverts - detritus weight

Model summary
```{r, echo=F}
summary(bact_detwt_lms[[1]])
```

Relative importance of each predictor
```{r, echo=F}
bact_detwt_relimp[[1]]
```

Stepwise AIC selected model
```{r, echo=F}
summary(bact_detwt_aicmod[[1]])
```

#### Detritus - detritus weight

Model summary
```{r, echo=F}
summary(bact_detwt_lms[[2]])
```

Relative importance of each predictor
```{r, echo=F}
bact_detwt_relimp[[2]]
```

Stepwise AIC selected model
```{r, echo=F}
summary(bact_detwt_aicmod[[2]])
```

#### Inverts - water volume

Model summary
```{r, echo=F}
summary(bact_wvol_lms[[1]])
```

Relative importance of each predictor
```{r, echo=F}
bact_wvol_relimp[[1]]
```

Stepwise AIC selected model
```{r, echo=F}
summary(bact_wvol_aicmod[[1]])
```

#### Water filters - water volume

Model summary
```{r, echo=F}
summary(bact_wvol_lms[[2]])
```

Relative importance of each predictor
```{r, echo=F}
bact_wvol_relimp[[2]]
```

Stepwise AIC selected model
```{r, echo=F}
summary(bact_wvol_aicmod[[2]])
```


## Significant predictors from each AIC-selected model

```{r, echo=F}
xx = data.frame(Locus = c(rep("Fungi", 4), rep("Bacteria", 4)),
           Compartment = rep(c("Inverts", "Detritus", "Inverts", "Water"), 2),
           Significant_predictors = c("None",
                                      "log(detritus_wt), Diameter, Invertebrate_number, pctC_detritus",
                                      "None",
                                      "pH, Temperature, Nitrate_water",
                                      
                                      "None",
                                      "log(detritus_wt), Diameter, Height, Complexity, Invertebrate_number, pctC_detritus",
                                      "log(water_volume), pH (marginal)",
                                      "Diameter, Height (marginal), Complexity (marginal), pH (marginal)")
           
           )

rmarkdown::paged_table(xx)
```

## relative importance

```{r}
fff <- lapply(fung_detwt_relimp, function(x) x$lmg)

fff2 <- lapply(fung_wvol_relimp, function(x) x$lmg)

fdw <- cbind(as.data.frame(fff[[1]]), as.data.frame(fff[[2]]))

fwv <- cbind(as.data.frame(fff2[[1]]), as.data.frame(fff2[[2]]))


bbb <- lapply(bact_detwt_relimp, function(x) x$lmg)

bbb2 <- lapply(bact_wvol_relimp, function(x) x$lmg)

bdw <- cbind(as.data.frame(bbb[[1]]), as.data.frame(bbb[[2]]))

bwv <- cbind(as.data.frame(bbb2[[1]]), as.data.frame(bbb2[[2]]))


all_dw <- cbind(fdw, bdw)
names(all_dw) <- c("Fungi, Invertebrates", "Fungi, Detritus", "Bacteria, Invertebrates", "Bacteria, Detritus")
hm <- as.data.frame(t(all_dw))

all_wv <- cbind(fwv, bwv)
names(all_wv) <- c("Fungi, Invertebrates", "Fungi, Water", "Bacteria, Invertebrates", "Bacteria, Water")

write.csv(all_dw, "../outputs/for_manuscript/detritus_weight_relimp.csv")
write.csv(all_wv, "../outputs/for_manuscript/water_volume_relimp.csv")

```

Stargazer table making

```{r}
dw_list <- list(fung_detwt_aicmod[[2]],
           fung_detwt_aicmod[[1]],
           bact_detwt_aicmod[[2]],
           bact_detwt_aicmod[[1]])

stargazer(dw_list,
          type = "html",
          digits = 3,
          header = FALSE,
          title= "Regression Results",
          out = "detritus_wt_mult_regress.htm",
          report=('vc*p')
         )


wv_list <- list(fung_wvol_aicmod[[2]],
           fung_wvol_aicmod[[1]],
           bact_wvol_aicmod[[2]],
           bact_wvol_aicmod[[1]])

stargazer(wv_list,
          type = "html",
          digits = 3,
          header = FALSE,
          title= "Regression Results",
          out = "water_vol_mult_regress.htm",
          report=('vc*p')
         )
```


